---
title: '**Chapter 6 Lab**'
author: "Jeremy Selva"
subtitle: This document was prepared on `r format(Sys.Date())`.
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: no
    code_folding: hide
    code_download: yes
    self_contained: true
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
bibliography: utils/bibliography.bib
csl: utils/f1000research.csl
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, warning=FALSE, message=TRUE, include=FALSE}
# Set up the environment

# Options relative to figure size
# 1.618 is the golden ratio
figheight <- 4
figwidth <- 4 * 1.618 

# General options
options(knitr.kable.NA = "",
        nsmall = 3,
        tidyverse.quiet = TRUE
        )
hook_output <- knitr::knit_hooks$get('output')

knitr::knit_hooks$set(
  output = function(x, options) {
    if (!is.null(options$max.height)) {
      options$attr.output <- c(options$attr.output,
                               sprintf('style="max-height: %s;"', options$max.height))
    }
    hook_output(x, options)
    }
  )

# Chunk options (see https://yihui.org/knitr/options/#chunk_options)
knitr::opts_chunk$set(
  comment = ">",  # The prefix to be added before each line of the text output.
  dpi = 600,
  fig.path = "img/",
  fig.height = figheight,
  fig.width = figwidth,
  fig.align = "center",
  # See https://community.rstudio.com/t/centering-images-in-blogdown-post/20962
  # to learn how to center images
  # See https://bookdown.org/yihui/rmarkdown-cookbook/opts-tidy.html
  # See https://www.zotero.org/styles for citation style respository
  tidy='styler',
  tidy.opts=list(strict=TRUE)
)
```

```{r warning=FALSE, message=FALSE, results='asis'}
library(leaps)
library(tibble)
library(magrittr)
library(dplyr)
library(tidyr)
library(report)
library(ISLR)

library(reactable)

library(parsnip)
library(glmnet)
library(rsample)
library(recipes)
library(tune)
library(workflows)
library(dials)

summary(report::report(sessionInfo()))
```

# Subset Selection Methods

## Best Subset Selection

```{r, echo=FALSE}
set.seed(1234)
```

We will be using the `Hitters` data set from the `ISLR` package. We wish
to predict the baseball players `Salary` based on several different
characteristics which are included in the data set.

Remove all rows with missing data from that column.

```{r, message=FALSE}

Hitters <- as_tibble(ISLR::Hitters) %>%
  tidyr::drop_na()
```

Use `regsubsets()` function to performs best subset selection.

An asterisk indicates that a given variable is included in the
corresponding model. For instance, this output indicates that the best
two-variable model contains only Hits and CRBI.

```{r, message = FALSE, max.height = '150px'}

regfit.full <- leaps::regsubsets(
  x = Salary ~ ., 
  data = Hitters,
  nvmax = 8 #default
  )

summary(regfit.full)
```

Here we fit up to a 19-variable model.

```{r, message=FALSE}

regfit.full <- leaps::regsubsets(
  x = Salary ~ ., 
  data = Hitters,
  nvmax = 19 #default
  )
```

The `summary()` function also returns $R^2$, $RSS$, adjusted $R^2$,
$C_p$, and $BIC$. We can examine these to try to select the best overall
model.

```{r, message=FALSE}

reg.summary <- regfit.full %>%
  summary()

names(reg.summary)
```

For instance, we see that the $R^2$ statistic increases from 32 %, when
only one variable is included in the model, to almost 55 %, when all
variables are included. As expected, the $R^2$ statistic increases
monotonically as more variables are included.

```{r, message=FALSE}

reg.summary$rsq
```

The `regsubsets()` function has a built-in `plot()` command which can be
used to display the selected variables for the best model with a given
number of predictors, ranked according to the $BIC$, $C_p$, adjusted
$R^2$, or $AIC$. To find out more about this function, type
`?plot.regsubsets`.

The top row of each plot contains a black square for each variable
selected according to the optimal model associated with that statistic.
For instance, we see that several models share a $BIC$ close to $−150$.
However, the model with the lowest $BIC$ is the six-variable model that
contains only `AtBat`, `Hits`, `Walks`, `CRBI`, `DivisionW`, and
`PutOuts`.

```{r, message=FALSE}
plot(regfit.full , scale = "bic")
```

We can use the `coef()` function to see the coefficient estimates
associated with this model.

```{r, message=FALSE}
coef(regfit.full , 6)
```

## Forward and Backward Stepwise Selection

We can also use the `regsubsets()` function to perform forward stepwise
selection using the argument `method = "forward"`

```{r, message=FALSE, max.height='150px'}
regfit.fwd <- leaps::regsubsets(
  x = Salary ~ ., 
  data = Hitters ,
  nvmax = 19, 
  method = "forward"
)

summary(regfit.fwd)
```

To perform backward stepwise selection, use the argument
`method = "backward"`

```{r, message=FALSE, max.height='150px'}
regfit.bwd <- leaps::regsubsets(
  x = Salary ~ ., 
  data = Hitters ,
  nvmax = 19, 
  method = "backward"
)

summary(regfit.bwd)
```

For this data, the best seven-variable models identified by forward
stepwise selection, backward stepwise selection, and best subset
selection are different.

```{r, message=FALSE}

coef(regfit.full,7)
coef(regfit.fwd,7)
coef(regfit.bwd,7)

```

## Choosing Among Models Using the Validation-Set Approach

We split the samples into training set and a test set

```{r, message=FALSE}

set.seed (1)

Hitters <- Hitters %>%
  dplyr::mutate(
    isTrainingSet = sample(x = c(TRUE , FALSE),
                           size = nrow(Hitters),
                           replace = TRUE)
  ) %>%
  dplyr::relocate(.data[["isTrainingSet"]])

train <- Hitters %>%
  dplyr::filter(.data[["isTrainingSet"]] == TRUE) %>%
  dplyr::select(-c("isTrainingSet"))

test <- Hitters %>%
  dplyr::filter(.data[["isTrainingSet"]]== FALSE) %>%
  dplyr::select(-c("isTrainingSet"))

```

Now, we apply `regsubsets()` to the training set in order to perform
best subset selection and compute the the validation set error for the
best model of each model size.

```{r, message=FALSE}

regfit.best <- leaps::regsubsets(Salary ~ ., data = train, nvmax = 19)

test.mat <- stats::model.matrix(Salary ~ ., data = test)

val.errors <- rep(NA, 19)

for (i in 1:19) {
  coefi <- stats::coef(regfit.best , id = i)
  pred <- test.mat[, names(coefi)] %*% coefi
  val.errors[i] <- mean (( test$Salary - pred)^2)
}

val.errors
```

We find that the best model is the one that contains seven variables.

```{r, message=FALSE}
which.min(val.errors)

coef(regfit.best , 7)
```

However, the best seven-variable model on the full data set has a
different set of variables than the best seven-variable model on the
training set.

```{r, message=FALSE}
regfit.best <- leaps::regsubsets(Salary ~ ., data = Hitters , nvmax = 19)
coef(regfit.best , 7)
```

## Choosing Among Models Using Cross-Validation

We now try to choose among the models of different sizes using cross
validation. Create a vector that allocates each observation to one of k
= 10 folds.

```{r, message=FALSE}
set.seed (1)

k <- 10
n <- nrow(Hitters)
folds <- sample(rep (1:k, length = n))

```

We write a for loop that performs cross-validation

```{r, message=FALSE}

# A predict method for regsubsets
predict.regsubsets <- function(object , newdata , id, ...) {
  form <- as.formula(object$call [[2]])
  mat <- model.matrix(form , newdata)
  coefi <- coef(object , id = id)
  xvars <- names(coefi)
  mat[, xvars] %*% coefi
  }

cv.errors <- matrix(NA, k, 19,
                    dimnames = list(NULL , paste (1:19)))
for (j in 1:k) {
  best.fit <- leaps::regsubsets(Salary ~ .,
                                data = Hitters[folds != j,],
                                nvmax = 19)
  for (i in 1:19) {
    pred <- stats::predict(best.fit, Hitters[folds == j, ], id = i)
    cv.errors[j, i] <- mean(( Hitters$Salary[folds == j] - pred)^2)
  }
  }

```

This has given us a $10$ by $19$ matrix, of which the $(j, i)$th element
corresponds to the test $MSE$ for the $j$th cross-validation fold for
the best $i$-variable model.

We use the `apply()` function to average over the columns of this
`apply()` matrix in order to obtain a vector for which the $i$th element
is the cross validation error for the $i$-variable model.

```{r, message=FALSE}

mean.cv.errors <- apply(cv.errors , 2, mean)

plot(mean.cv.errors , type = "b")

```

We see that cross-validation selects a 10-variable model.

```{r, message=FALSE}
regfit.best <- leaps::regsubsets(Salary ~ ., data = Hitters , nvmax = 19)
coef(regfit.best , 10)
```

# Ridge Regression and the Lasso

## Ridge Regression

Set mixture to 0 to indicate Ridge Regression. For now, we set
penalty/lambda as 0

```{r, message=FALSE}
Hitters <- as_tibble(ISLR::Hitters) %>%
  tidyr::drop_na()

ridge_spec <- parsnip::linear_reg(mixture = 0,
                                  penalty = 0) %>%
  parsnip::set_mode("regression") %>%
  parsnip::set_engine("glmnet")

ridge_spec %>%
  parsnip::translate()
```

Once the specification is created we can fit it to our data. We will use
all the predictors.

```{r, message=FALSE}
set.seed(1)
ridge_fit <- ridge_spec %>%
  parsnip::fit(formula = Salary ~ ., 
               data = Hitters)
```

We can see the parameter estimate for different values of
penalty/lambda. Notice how the estimates are decreasing when the amount
of penalty goes up

```{r, message=FALSE, max.height='150px'}
parsnip::tidy(ridge_fit)
parsnip::tidy(ridge_fit, penalty = 705)
parsnip::tidy(ridge_fit, penalty = 11498)
```

We can visualize how the magnitude of the coefficients are being
regularized towards zero as the penalty goes up.

```{r, message=FALSE}
# Arrange figure margin
par(mar=c(5,5,5,1))
ridge_fit %>%
  parsnip::extract_fit_engine() %>%
  plot(xvar = "lambda")
```

It would be nice if we could find the "best" value of the
penalty/lambda. We can do this using the `tune::tune_grid()`

Need we need three things in order to use the `tune::tune_grid()`

-   a `resample` object containing the resamples the `workflow` should be
    fitted within,
    
-   a `workflow` object containing the model and preprocessor,and

-   a tibble `penalty_grid` containing the parameter values to be evaluated.

First, we create a 10-fold cross-validation data set from the training
data set.

### Create the resample object

```{r, message=FALSE}

Hitters_split <- rsample::initial_split(Hitters, strata = "Salary")

Hitters_train <- rsample::training(Hitters_split)
Hitters_test <- rsample::testing(Hitters_split)

Hitters_fold <- rsample::vfold_cv(Hitters_train, v = 10)

```


### Create the preprocessor

We use the `recipes` package to create the preporcessing steps. However, the order of the preprocessing step is actually important. See the [Ordering of steps vignette](https://cran.r-project.org/web/packages/recipes/vignettes/Ordering.html) 

We create a basic recipe.

```{r, message=FALSE}

ridge_recipe <- recipes::recipe(formula = Salary ~ .,
                                data = Hitters_train)

rec <- recipes::prep(x = ridge_recipe, training = Hitters_train)
  
rec %>%
  summary() %>%
  reactable::reactable(defaultPageSize = 5)

rec %>%
  recipes::bake(new_data = Hitters_train) %>%
  reactable::reactable(defaultPageSize = 5)

```

Do preprocessing on the factor variables.

```{r, message=FALSE}

ridge_recipe <- ridge_recipe %>%
  # Step Novel is important.
  # If test set has a new factor not found in training, it will be treated as "new" and not NA
  # This will prevent the model from giving an error
  # https://blog.datascienceheroes.com/how-to-use-recipes-package-for-one-hot-encoding/
  recipes::step_novel(recipes::all_nominal_predictors()) %>%
  # Create Dummy variables
  recipes::step_dummy(recipes::all_nominal_predictors())

rec <- recipes::prep(x = ridge_recipe, training = Hitters_train)
  
rec %>%
  summary() %>%
  reactable::reactable(defaultPageSize = 5)

rec %>%
  recipes::bake(new_data = Hitters_train) %>%
  reactable::reactable(defaultPageSize = 5)

```

Do normalisation step.

```{r, message=FALSE}

ridge_recipe <- ridge_recipe %>%
  # Remove predictors with zero variation
  recipes::step_zv(recipes::all_predictors()) %>%
  # Standardise all variable
  recipes::step_normalize(recipes::all_predictors())

rec <- recipes::prep(x = ridge_recipe, training = Hitters_train)
  
rec %>%
  recipes::bake(new_data = Hitters_train) %>%
  reactable::reactable(defaultPageSize = 5)


```

### Specify the model

The model specification will look very similar to what we have seen earlier, but we will set `penalty = tune::tune()`. This tells `tune::tune_grid()` that the penalty parameter should be tuned.

```{r, message=FALSE}

ridge_spec <- 
  parsnip::linear_reg(penalty = tune::tune(), mixture = 0) %>% 
  parsnip::set_mode("regression") %>% 
  parsnip::set_engine("glmnet")


```

### Create the workflow

```{r, message=FALSE}

ridge_workflow <-  workflows::workflow() %>% 
  workflows::add_recipe(ridge_recipe) %>% 
  workflows::add_model(ridge_spec)

ridge_workflow
```

### Create the penalty/lambda grid

```{r, message=FALSE}

penalty_grid <- dials::grid_regular(dials::penalty(range = c(-5, 5)), levels = 50)

penalty_grid  %>% 
  reactable::reactable(defaultPageSize = 5)
```
